Comparison of several stochastic and deterministic derivative-free global optimization algorithms\thanks{The study was supported by the Russian Science Foundation, project No 16-11-10150.}

### 2019-02-11 2019/MOTOR_Sovrasov/motor_paper.tex
MOTOR: update conclusions

\begin{abstract}
In this paper popular sequential open-source solvers are compared against
Globalizer solver, which is developed at the Lobachevsky State University.
The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
competitive performance with other similar solvers. The comparison is done on several sets of
challenging multi-extremal benchmark functions. Also this work considers a method of
heuristic hyperparameters control for Globalizer allowing to reduce amount of
initial tuning before optimization. The proposed scheme allows rapidly increase convergence speed
of Globalizer by switching between ``local'' and ``global'' search phases in runtime.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box
optimization}
\end{abstract}
%
\section{Introduction}

@@

В данной работе рассмотрены open-source реализации восьми различных методов глобальной оптимизации, представленные
в библиотеке NLOpt\cite{nlopt} и пакете SciPY\cite{scipy}.
Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.

@@

\subsubsection{Hyperparameters tuning}
Параметр $r$ из (\ref{eq:step2}) непосредственно влияет на глобальную сходимость AGS (see \cite{strSergGO}, Chapter 8):
при достаточно большом значении $r$ метод гарантированно сходится ко всем глобальным минимумам целевой функции.
В то же время, согласно (\ref{eq:step3_2}), при бесконечно большом значении $r$ AGS превращается в полный перебор
по равномерной сетке.

@@

В указанном алгоритме $r_{min} < r_{max}$, $q > 1$. Вместо одного параметра $r$ теперь
необходимо выбирать 3, однако, согласно численным экспериментам, сделать это проще, чем найти оптимальное значение $r$.
Интуитивно практическую эффективность предложенной схемы можно объяснить тем, что теперь работа метода происходит в
двух режимах: глобальный поиск при $r=r_{max}$ и локальный при $r=r_{min}$. Если во время фазы глобального посика
метод приблизился к глобальному минимуму, то во время следующей фазы оценка глобального минимума будет быстро уточнена.
Если двух фаз не хватило, то процесс продолжается. Таким образом выбирается более оптимальный trade-off between
exploration and exploitation. Далее будем обозначать метод, использующий описанную схему как AGS-AR.

@@

\begin{table}
\begin{center}
\caption{Number of test optimization problems solved by methods}
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
  \hline
  GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
  \hline
  GKLS 2d Hard   &  100 & 100 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
  \hline
  GKLS 3d Simple &  100 & 97  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
  \hline
  GKLS 3d Hard   &  100  & 99   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
  \hline
  GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
  \hline
  GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
  \hline
  GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
  \hline
  GKLS 5d Hard   &  97  & 99  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
  \hline
  \end{tabular}
  \label{tab:solved}
\end{center}
\end{table}

@@

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item the proposed modification of the stock AGS substantially outperforms it on the most complicated problems
  from the generated test set and allows to pay less attention to initial hyperparameter tuning.
  \item AGS-AR method  has demonstrated the convergence
speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
\end{itemize}

>>>>

\begin{abstract}
In this paper popular sequential open-source solvers are compared against
Globalizer solver, which is developed at the Lobachevsky State University.
The Globalizer is designed to solve problems with black-box objectives satisfying the Lipschitz condition and shows
competitive performance with other similar solvers. The comparison is done on several sets of
challenging multi-extremal benchmark functions. Also this work considers a method of
heuristic hyperparameters control for Globalizer allowing to reduce amount of
initial tuning before optimization. The proposed scheme allows rapidly increase convergence speed
of Globalizer by switching between ``local'' and ``global'' search phases in runtime.
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{deterministic global optimization $\cdot$ stochastic global optimization
  $\cdot$ parallel numerical methods $\cdot$ derivative-free algorithms $\cdot$ black-box
optimization}
\end{abstract}
%
\section{Introduction}
\begin{Russian}

@@

В данной работе рассмотрены open-source реализации восьми различных методов глобальной оптимизации, представленные
в библиотеке NLOpt\cite{nlopt} и пакете SciPY\cite{scipy}.
Все алгоритмы были протестированы на наборе из 900 существенно многоэкстремальных функций, который был сгенерирован с
помощью специализированных генераторов задач \cite{Gaviano2003, grishaginClass}.
\end{Russian}

@@

\subsubsection{Hyperparameters tuning}

\begin{Russian}
Параметр $r$ из (\ref{eq:step2}) непосредственно влияет на глобальную сходимость AGS (see \cite{strSergGO}, Chapter 8):
при достаточно большом значении $r$ метод гарантированно сходится ко всем глобальным минимумам целевой функции.
В то же время, согласно (\ref{eq:step3_2}), при бесконечно большом значении $r$ AGS превращается в полный перебор
по равномерной сетке.

@@

В указанном алгоритме $r_{min} < r_{max}$, $q > 1$. Вместо одного параметра $r$ теперь
необходимо выбирать 3, однако, согласно численным экспериментам, сделать это проще, чем найти оптимальное значение $r$.
Интуитивно практическую эффективность предложенной схемы можно объяснить тем, что теперь работа метода происходит в
двух режимах: глобальный поиск при $r=r_{max}$ и локальный при $r=r_{min}$. Если во время фазы глобального посика
метод приблизился к глобальному минимуму, то во время следующей фазы оценка глобального минимума будет быстро уточнена.
Если двух фаз не хватило, то процесс продолжается. Таким образом выбирается более оптимальный trade-off between
exploration and exploitation. Далее будем обозначать метод, использующий описанную схему как AGS-AR.
\end{Russian}

@@

\begin{table}
\begin{center}
\caption{Number of test optimization problems solved by methods}
  \begin{tabular}{|l|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|{c}|}
    \hline
    & AGS & AGS-AR & CRS & DIRECT & DIRECT\(l\) & MLSL & SDA & DE & StoGO \\
  \hline
  \(F_{GR}\)     &  100 & 100 & 76  & 100 & 100 & 97  & 96  & 96  & 67\\
  \hline
  GKLS 2d Simple &  100 & 100 & 85  & 100 & 100 & 100 & 100 & 98  & 90\\
  \hline
  GKLS 2d Hard   &  100 & 97 & 74  & 100 & 100 & 100 & 93  & 85  & 77 \\
  \hline
  GKLS 3d Simple &  100 & 100  & 75  & 100 & 100 & 100 & 89  & 86  & 44 \\
  \hline
  GKLS 3d Hard   &  100  & 100   & 72   & 100  & 99   & 100  & 88   & 77   & 43 \\
  \hline
  GKLS 4d Simple &  100 & 100 & 46  & 100 & 100 & 94  & 78  & 59  & 16 \\
  \hline
  GKLS 4d Hard   &  100 & 100 & 47  & 99  & 97  & 94  & 72  & 32  & 10  \\
  \hline
  GKLS 5d Simple &  100 & 100 & 68  & 100 & 100 & 98  & 100 & 77  & 9  \\
  \hline
  GKLS 5d Hard   &  97  & 100  & 42  & 100 & 90  & 79  & 84  & 48  & 8 \\
  \hline
  \end{tabular}
  \label{tab:solved}
\end{center}
\end{table}

@@

In the present paper, several sequential and parallel global optimization algorithms were considered.
A comparison of efficiencies of these ones has been done on a set of test problems. The results allow
making the following conclusions:
\begin{itemize}
  \item the proposed modification of the stock AGS, AGS-AR allows to pay less attention to initial hyperparameter tuning and
  performs on-par with properly tuned AGS.
  \item AGS-AR method has demonstrated the convergence
  speed and reliability at the level of DIRECT and exceeds many other algorithms, the open-source
  implementations of which are available;
  \item the stochastic optimization methods inferior to the deterministic ones in the convergence
speed and in reliability. It is manifested especially strongly on more complex multiextremal
problems;
\end{itemize}
